---
title: "Project 3: Data Analysis Pipeline"
author: "Marc R Radke"
date: "18 March 2021"
output: html_document

---

<!--- Begin styling code. --->

```{=html}
<style type="text/css">
/* Whole document: */
body{
  font-family: Cambria, Georgia, Palatino, serif;
  font-size: 12pt;
}
h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author {
  font-size: 18px;
  text-align: center;
}
h4.date {
  font-size: 18px;
  text-align: center;
}
</style>
```
<!--- End styling code. --->

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To demonstrate the data analysis pipeline, we will use the random forest cross-validation function with the `my_penguins` data. First, we'll load the data and source our function:

```{r, message = FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
# Load data
my_penguins <- read.csv("../Data/my_penguins.csv")
# Trim penguins dataset to only variables we want  
pen <- my_penguins[,3:6] %>% tidyr::drop_na()
# Source my_rf_cv.R
source("../Code/my_rf_cv.R")
```

We will predict `body_mass_g` using  `bill_length_mm`, `bill_depth_mm`, and `flipper_length_mm` and iterate through `k = 2`, `k = 5`, and `k = 10` (where `k` is the number of folds in our cross-validation) 30 times each to analyze the spread of the average MSE (mean squared error).

```{r, cache = TRUE}
reps <- 30
k_vals <- c(2, 5, 10)
# Create empty output dataframe
out <- data.frame("k" = rep(NA, reps * length(k_vals)),
                  "mse" = rep(NA, reps * length(k_vals)))

# Iterate through each k and each repetition and record the MSE
start <- 0
for (k in k_vals) {
  for (i in 1:reps) {
    out[start + i, 1] <- paste0("k_", k)
    out[start + i, 2] <- my_rf_cv(k)
  }
start <- start + reps
}
# Save data in Results folder
write.csv(out, file = "../Output/Results/rf_simulations.csv", row.names = FALSE)
```

Next, we'll plot the MSE for each `k` and summarize the data:

```{r, message = FALSE, fig.width = 7, fig.height = 5, fig.align = "center"}
# Make k column a factor and relevel
out$k <- factor(out$k, levels = c("k_2", "k_5", "k_10"))

# Create boxplots for each value of k
rf_plot <- ggplot(data = out, aes(y = mse, x = k)) +
  geom_boxplot(fill = "#ff6961") +
  scale_x_discrete(labels = c(2, 5, 10)) +
  labs(title = "Mean Squared Error For Each k",
       y = "MSE") +
  theme_dark() +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 1, 0.5, 0.5, "cm"))
# Save plot to Figures folder
ggsave("../Output/Figures/MSE_boxplot.png")
rf_plot
```

```{r}
# Group by k and summarize: mean and sd
summary <- out %>%
  # Change k value strings to numbers so table looks nicer
  dplyr::mutate(k = rep(c(2,5,10), each = 30)) %>%
  dplyr::group_by(k) %>%
  dplyr::summarize(mse_mean = mean(mse),
            mse_sd = sd(mse))
# assign colnames for printing
colnames(summary) <- c("k", "Mean MSE", "SD MSE")
# Save table to Results folder
saveRDS(summary, file = "../Output/Results/summary.rds")
# Format and print table
kable_styling(kable(summary), full_width = FALSE, position = "center")
```

As `k` increases, the mean and standard deviation of the MSE tends to decrease, possibly because more folds means that more training data is used to fit the model each time.
